### 2011-06-02 thesis/novelty-intro.tex
Changes on review of novelty detection.

This chapter presents a statis

Laying down base concepts that are then used in \autoref{chap:novelty} to
implement novelty detection on the semantic spatial representation.

It start by presenting a brief overview on novelty detection.

It introduces novelty as a

Shows novelty has an associated ordering.
Shows novelty has a threshold function.

At last a practical example using semantic data and probabilistic graphical
models is presented: it shows how to use the introduced ratio to obtain a
novelty detection system and analyses the performance impact by obtaining more
data and by using an approximation on the unconditional probability.

\begin{itemize}
\item Introduction
\item History
\item Approaches to Novelty Detection
\end{itemize}

\section{Novelty Detection Review and Related Work}
% Introduce novelty detection
Novelty detection deals with detecting that a new input was generated by a class
other than the ones the system knows about~\cite{markou2003novelty}.
It is harder than classification as only positive samples of a class are available
rendering normal classification methods invalid.

Adding capabilities to an agent to detect novel samples allows to increase its
reliability. Novelty signal can then be interpreted by the system to proceed
with more caution as its knowledge does not correctly describe reality.

@@

\section{Novelty Detection as a Threshold}
When dealing with a statistic point of view, noisy data or unstable features,
a decision to classify a sample $n$ as novel has an associated error probability
$P(\overline{novel}|n)$.
Accordingly, on a system that classifies a set of inputs $N$ as novel, the
probability of incorrectly classifying a sample as novel is given by the
equation:

>>>>

This chapter presents concepts that are used as base when adding novelty
detection capabilities on the semantic representation in \autoref{chap:novelty}.
It starts by providing the reader with a brief overview on novelty detection
techniques and related works.

After it shows that an optimal novelty detection system can be implemented by
thresholding on a order-relation defined over the inputs.
And that an equivalent order-relation can be imposed by a ratio between
conditional and unconditional probabilities.
Some discussion on how to interpret the meaning of those factors is also
presented.

At last a practical example using semantic data and probabilistic graphical
models is presented: it shows how to use the introduced ratio to obtain a
novelty detection system and analyses the performance impact by increasing
the amount of sensed data and by using an approximation on the unconditional
probability.

\section{Novelty Detection Review and Related Work}
Novelty detection, also called outlier or anomaly detection, is the
classification problem related with identification of new or unknown data
patterns that the system is not aware of~\cite{markou2003novelty}.

The ability to identify novel cases is crucial in any autonomous system
that is deployed to unknown or to uncontrolled environments, as it gives the
system the ability to detect something is not conforming to it knowledge and
therefore it should proceed with caution.
It has several applications such as fault detection, intrusion detection,
detection of masses in mammograms, hand written digit recognition and many
others.
% TODO: list usage of novelty detection together with references to papers
% with them.

It is on the nature of unknown environments and/or in the infeasibility of
training the system on all possible object classes that lies the complexity of
novelty detection. It is only possible to obtain samples representing positives
examples of known cases. The lack of negative examples renders normal
classification methods unusable.

In order to be usable in real world applications, novelty detection methods
have to overcome a series of obstacles:
be able to generalize while still detecting novelty,
be resistant to noisy features,
ability to scale in feature dimension,
deal with multiple classes and performing detection efficiently (i.e. many
autonomous systems require real-time or close to real-time performance).

\subsection{Review}
Many approaches have been made based on statistical methods. Those often model
the training data with statistic distributions and then perform tests whether a
given sample comes from the same distribution.

Parametric approaches: probabilistic /gmm approaches, hidden markov models,
hypothesis testing

Non-Parametric: knn Approachs, parzen density estimation, string matching,
clustering.

One of the simplest approach for novelty detection is pose the problem as a
probability density estimation problem and use the expected probability of a
given input to trigger the case as novel.

Examples of such approaches are Gaussian Mixtures Models and Parzen-window estimators.
In order to reduce the input size, those approaches often employ
dimension-reduction techniques such as \gls{PCA}.

This approach is used in \cite{bishop1994novelty} where \emph{novelty detection}
is added to a \emph{neural network} system. On it the density estimation on the
input is calculated using a \emph{Parzen window} over training-data and that
value is used as a novelty measure. This way they detect novel data that their
network has not seen before and the output of their system is then not expected
to be reliable.

A slightly different approach is done by one-class \gls{SVM} approaches that
try to distinguish the class by separating the traing set from all the other
points in the input space. They try to achieve that by enclosing the training
set by some structure (i.e. an hyper-sphere)\cite{bennett2000support}.
Those approaches have been made in line with Vapnik principle of not solving
something hard. As although having access to a perfect probability distribution
of the input would solve the problem, creating such a function is harder than
simply creating a boundary between known data and novel data
\cite{scholkopf2000support}.

Error reconstruction methods have also been used for novelty detection.
They use the assumption that the class to be defined lies on a manifold embedded
on the input space of higher dimensions. By using dimensional reduction
techniques, they try to defined that manifold and calculate how far a new sample
is from it.
One of the most common methods used for that is
\gls{K-PCA}~\citep{scholkopf1997kernel}, which uses the kernel-trick to extend
\gls{PCA} and perform a nonlinear dimensionality reduction on the input.
This technique has been successfully used in novelty detection by
\cite{Hoffmann2007863}.

\subsection{Related Work}
More in line with novelty detection in scene categorization and semantic
structures, the work of PLISS\dots

\section{Novelty Detection as a Threshold}

@@

When dealing with a statistic point of view, noisy data or unstable features,
a decision to classify a sample $n$ as novel has an associated error probability
$P(\overline{novel}|n)$.
Accordingly, on a system that classifies a set of inputs $N$ as novel, the
probability of incorrectly classifying a sample as novel is given by the
equation:

### 2011-06-02 thesis/novelty.tex
Changes on review of novelty detection.

To handle those issues a fully probabilistic approach should be taken
instead of a deterministic.
That can be done by extending all the variables on the factor graph representing
the semantic map with an additional class \emph{novel} and extending all the
factors with information on the unconditional distribution.

>>>>

To handle those issues a fully probabilistic approach should be taken
instead of a deterministic.
That can be done by extending all the variables on the factor graph representing
the semantic map with an additional state \emph{novel} and extending all the
factors with information on the unconditional distribution.
